---
title: "Yang_data_analysis"
output: word_document
---

Load required packages
```{r}
library(ggplot2)
library(dplyr)
library(broom)
```

load data. 
```{r}
baseJ_KO_gene_change <- readRDS("../../data/processed_data/baseJ_KO_gene_change.rds")
H3V_KO_gene_change <- readRDS("../../data/processed_data/H3V_KO_gene_change.rds")
double_KO_gene_change <- readRDS("../../data/processed_data/double_KO_gene_change.rds") 
life_stage_gene_expression <- readRDS("../../data/processed_data/life_stage_gene_expression.rds")
```

First, I ask if genes whose expressions were affected by base J knockout are related to certain stages of T. brucei life cycle. 
```{r}
baseJ_KO_gene_change
life_stage_gene_expression
BaseJ_related_genes <- baseJ_KO_gene_change$name
cluster1 <- life_stage_gene_expression %>%
  filter(`Gene ID` %in% BaseJ_related_genes)
cluster1
```
To see if any of these genes are up-regulationed by a certain cell cycle stage, I need to compare the ratios of expression values between different life stages, as presented in the last columns in cluster1 table. If the ratio for a specific gene is greater than 2, then I consider that gene to be up-regulated.  

```{r}
cluster1 %>%
  filter(`FC St/Sl` > 2) %>%
  select(`Gene ID`) %>%
  count()
cluster1 %>%
  filter(`FC St/Early` > 2) %>%
  select(`Gene ID`) %>%
  count()
cluster1 %>%
  filter(`FC St/Late` > 2) %>%
  select(`Gene ID`) %>%
  count()
cluster1 %>%
   filter(`FC Early/Late` > 2) %>%
  select(`Gene ID`) %>%
  count()
number_genes <- c(10, 10, 12, 5)
genes_affected_baseJ_KO <- matrix(number_genes, ncol = 1, byrow = TRUE)
rownames(genes_affected_baseJ_KO) <- c("stumpy", "slender", "early", "late")
genes_affected_baseJ_KO <- as.table(genes_affected_baseJ_KO)
genes_affected_baseJ_KO <- as.data.frame(genes_affected_baseJ_KO)
genes_affected_baseJ_KO %>%
  ggplot(aes(Var1, Freq)) + geom_point()
```

From this plot, it seems that base J KO had the least impact on gene expression of late procyclic form of T. brucei. 









